---
title: "Data_visualization"
author: "Adelaide Robinson"
date: '2022-08-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(gapminder)
library (sf)

```

```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

```{r}
lizards |> ggplot(aes(x = total_length, y = weight)) + 
  geom_point()
# geom_point can change to any compatable type
```

```{r}
lizards |> ggplot(aes(x = total_length)) + 
  geom_histogram(color = "orange",
                 size = 0.3,
                 fill = "purple",
                 linetype = "dotted"
                 )
#changes aesthetic not based on variable, don't put w/in aes

```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight))+
  geom_point(
  shape = 22,
  color = "cyan 4",
  fill= "yellow",
  size = 4,
  alpha = 0.4,
)
```

```{r}
ggplot(data = lizards,
       aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name, size = total_length),
             alpha = 0.5) +
  facet_wrap(~common_name,
             scales = "free") #change the scales
  theme_minimal()

```

total counts of lizards by common name

```{r}

new_lizard <- lizards |> 
  group_by(common_name) |> 
  summarize(
    counts = n()
  )
# do it using counts
lizard_counts <- lizards |> 
  count(common_name)

```

# plot it

#default is alphabetical order

```{r}
ggplot(data = lizard_counts, 
       aes(y = fct_reorder(common_name, n),# reorder common names by number
           x = n)) +
  geom_col()
#we want to change the order
```

create a scatterplot, faceted by species, for total length vs weight the color and shape of points should differ based on the species change the theme, make each faceted panel have flexible axes limits(scale is free)

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name, shape = common_name), 
             fill = "black",
             size = 1,
             alpha = 0.6) +
  theme_minimal() +
  labs(x = "Total length (mm)",
       y = "Weight (g)",
       color = "Lizard species") +
  facet_wrap(~common_name, scales = "free")
```

### Fun with scales

```{r}

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradient(low = "purple", high ="orange")



ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradientn(
    colors = c("magenta", "cyan 4", "darkorchid3", "dodgerblue","green", "yellow"))

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_steps(low = "red", high = "black")

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_steps2(
    low = "purple",
    mid = "white",
    high = "black",
    midpoint = 50,
    breaks = c(15, 30, 45))



ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_stepsn(colors = c("orange",
                                "red",
                                "purple"),
                     breaks = seq(from = 0, to =300, by = 60))


```

you can create a package with a graph theme

# factor reorder

```{r}
lizards_median <- lizards |> 
  mutate(common_name = fct_reorder(common_name, total_length,
                                   .fun = median))
#create a new column, overwrites existing column, recast common name as a factor 

  
ggplot(data = lizards_median, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "nationalparkcolors::CraterLake")+
  theme_minimal()
  
view(paletteer::palettes_d_names) # look at color sets
```

```{r}

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) + 
  theme(panel.grid.major.x = element_line(color = "red"),
        panel.grid.minor.y = element_blank(), # leave it blank
        axis.title.x = element_text(color = "purple"),
        axis.text.y = element_text(color = "green"),
        text = element_text(size = ),
        panel.background = element_rect(color = "purple", fil = "yellow")) + annotate("text", x = 300, y = 50, label = "woooo", color = "green") +
  geom_vline(xintercept = 250, linetype = "dashed",
             color = "orange")
      
```

Repulsive labels w/ ggrepel

```{r}

wws_lizards <- lizards |> filter(common_name == "western whiptail",
                      site == "sand")

#bad labels
ggplot(data = wws_lizards, aes(x = total_length, y = weight)) +
  geom_point()+
  geom_text(aes(label = toe_num))

#dodge points
ggplot(data = wws_lizards, aes(x = total_length, y = weight)) +
  geom_point()+
  geom_text_repel(aes(label = toe_num))


gapminder |> 
  filter(year == 2002, continent == "Europe") |> 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country, size = 2))

```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  gghighlight(toe_num == 250, label_key = toe_num)+ #gets overwritten by next line
  gghighlight(weight >30, label_key = toe_num)
```

```{r}
jornada_veg <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml"))


ggplot() + 
  geom_sf(data = jornada_veg,
          aes(fill = Name), color = NA) +
  theme_minimal() +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys")
```

```{r}

lizard_counts <- lizards |> 
  mutate(date = lubridate::mdy(date)) |> 
  count(year = lubridate::year(date), common_name) |> 
  drop_na()

```

```{r}
ggplot(data = lizard_counts, aes(x = year, y = common_name)) +
  geom_tile(aes(fill= n), show.legend = FALSE)+
    geom_text(aes(label = n), color = "white", size =3) +
  scale_fill_gradientn(colors = c("navy", "red", "orange"))+
  theme_minimal()
```

### Beeswarm plot

```{r}
whiptails <- lizards |> 
  filter(common_name == "western whiptail") |> 
  drop_na(total_length, weight)

whiptail_bee <- ggplot(data = whiptails, aes(x = sex, y = weight)) +
  geom_beeswarm(size = 0.5) + geom_boxplot(fill = NA)

```

### marginal plot

```{r}
whiptail_plot <- whiptails |> 
  ggplot(aes( x = total_length, y = weight)) + geom_point(aes(color = sex)) +
  theme(legend.position = "bottom")


ggMarginal(whiptail_plot, type = "boxplot", groupColour = TRUE)
```

### patchwork

```{r}
(whiptail_bee + whiptail_plot / whiptail_bee) #side by side

(whiptail_bee / whiptail_plot) #columns



```
